Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-18 surfaces-write-up/surface-write-up.tex
added remark on tightness of bounds to intro

\noindent
with relations generated in degrees at most
\[
	\max \left(2 \sum_{i=0}^n \ell_i a_i, \frac{\max_{0\le i \le n}
	(	\bida_i)}{\deg(D)} + \sum_{i=0}^n \ell_i a_i \right).
\]
\end{restatable}

\noindent
In Section \ref{sec:hirz}, we shift our attention to Hirzebruch surfaces. Recall that for each $m \geq 0$ we can define the
Hirzebruch surface $F_m := \proj(\sym^\bullet \sco_{\bp^1} \oplus
\sco_{\bp^1}(m)),$ or as a quotient of an open subscheme of
$\ba^4$ by a 
weighted action, as is done in the
beginning of Section \ref{sec:hirz}.
Note that Hirzebruch surfaces are regular, normal,
and locally factorial, being $\bp^1$ bundles over
$\bp^1$, and therefore the natural map from Cartier
divisors to Weil divisors is an isomorphism.
In other words, all Cartier divisors arise as the zero locus
of a section of some line bundle. We implicitly use this in the
statement of the following theorem. The analogous fact for
projective space was also implicitly used in Theorems
~\ref{thm:proj-effective-intro} and ~\ref{thm:proj-generators-relations}.

>>>>

\noindent
with relations generated in degrees at most
\[
	\max \left(2 \sum_{i=0}^n \ell_i a_i, \frac{\max_{0\le i \le n}
	(	\bida_i)}{\deg(D)} + \sum_{i=0}^n \ell_i a_i \right).
\]
\end{restatable}
\begin{rem}\label{rem:proj-tight-bounds} Note that the bounds
	given in Theorem ~\ref{thm:proj-effective-intro}
are tight and are typically attained, as explained in
Remark ~\ref{rem:exact-effective-bounds}.
Additionally, the bounds given in Theorem 
~\ref{thm:proj-generators-relations} are asymptotically
tight to within a factor of two 
for the following class of divisors. Consider
a divisor 
$ D = \sum_{i=0}^n\frac{p_i}{2q_i}H_i \in \bq \otimes_\bq \di \bp^m$
such that $H_i$ are hyperplanes,  
$q_i$ are pairwise coprime integers, and $p_i$ are chosen
so that $\deg D = \frac{1}{2 \prod_{i=0}^n}q_i$. Further,
choose a linear subspace $\pi: \bp^1 \rightarrow \bp^m$  generically so that
$\pi^* D = \sum_{i=0}^n \frac{p_i}{2q_i}P_i$, 
where $P_i$ are distinct points in $\bp^1$. To choose such a map
$\pi$, we may need to assume that the base field is infinite.
Note that by the Remark immediately following Theorem 8 in 
\cite{dorney:canonical},
the given bounds on the generators and relations of
$R_{\pi^*D}$ are within
a factor of two of the degree of generation and relations of
$R_{\pi^*D}$. Finally, since restriction map
$R_D \rightarrow R_{\pi^*D}$ induced by the restriction maps
on cohomology $H^0(\bp^m, dD) \rightarrow H^0(\bp^1, \pi^*(dD))
\cong H^0(\bp^1, d\pi^* D)$ are surjective, we obtain that
the bounds for the generators and relations of $R_D$ also
agree with the degree of generation and relations 
to within a factor of two.
 \end{rem}

In Section \ref{sec:hirz}, we shift our attention to Hirzebruch surfaces. Recall that for each $m \geq 0$ we can define the
Hirzebruch surface $F_m := \proj(\sym^\bullet \sco_{\bp^1} \oplus
\sco_{\bp^1}(m)),$ or as a quotient of an open subscheme of
$\ba^4$ by a 
weighted action, as is done in the
beginning of Section \ref{sec:hirz}.
Note that Hirzebruch surfaces are regular, normal,
and locally factorial, being $\bp^1$ bundles over
$\bp^1$, and therefore the natural map from Cartier
divisors to Weil divisors is an isomorphism.
In other words, all Cartier divisors arise as the zero locus
of a section of some line bundle. We implicitly use this in the
statement of the following theorem. The analogous fact for
projective space was also implicitly used in Theorems
~\ref{thm:proj-effective-intro} and ~\ref{thm:proj-generators-relations}.
